---
title: "Homework 5: how do Sarracenia characteristics predict biomass?"
author: "Katie Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc-location: left
    code-fold: TRUE
    theme: yeti
execute:
  message: FALSE
  warning: FALSE
---

# Introduction

```{r packages}
# loading in necessary packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```


```{r read-data}
# reading in the data, filtering for clarity and relevancy
plant <- read_csv(here("data", "knb-lter-hfr.109.18","hf109-01-sarracenia.csv")) %>% 
  # make column names cleaner
  clean_names() %>% 
  # select columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```


# Methods
Species

a. Sampling methods (these should be summarized from the metadata and cited)

b. Description of data organization and processing (1-3 sentences)

c. We began by creating a visualization of the missing data in our entire dataset (Figure 1). 
Visualization of missing observations with an accompanying caption and discussion
(caption: 1-3 sentences, discussion: 1-3 sentences)

d. Visualization of Pearson’s correlation between variables with an accompanying
caption and discussion (caption: 1-3 sentences, discussion: 1-3 sentences)

e. Visualization of relationships between variables (not correlation) with an
accompanying caption and discussion (caption: 1-3 sentences, discussion: 1-3
sentences)

f. Creating the null and full model with accompanying discussion (1-3 sentences)
g. Visual and statistical assumption checks for the full model with accompanying
discussion (1-3 sentences)

h. Explanation of any transformation and justification (1-3 sentences)

i. Model construction with visual and statistical assumption checks for three
additional models with justification (i.e. why did you choose those predictor
variables?) (justification: 1-3 sentences per model)

j. Variance inflation factor check for the full model with accompanying discussion
(discussion: 1-3 sentences)

k. Model comparison using the null model, full model, and additional models from part
7g

```{r figure-1}
# visualize missing data 
gg_miss_var(plant)
```


# Results 
Species:

# Bibliography
