---
title: "Homework 5: how do Sarracenia characteristics predict biomass?"
author: "Katie Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc-location: left
    code_folding: true
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction


```{r packages}
# loading in necessary packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```


```{r read-data}
# reading in the data, filtering for clarity and relevancy
plant <- read_csv(here("data", "knb-lter-hfr.109.18","hf109-01-sarracenia.csv")) %>% 
  # make column names cleaner
  clean_names() %>% 
  # select columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```


# Methods
Species

a. Sampling methods (these should be summarized from the metadata and cited)

b. Description of data organization and processing (1-3 sentences)

c. We began by creating a visualization of the missing data in our entire dataset (Figure 1). 
Visualization of missing observations with an accompanying caption and discussion
(caption: 1-3 sentences, discussion: 1-3 sentences)

d. In order to determine the relationships between numerical variables in our dataset, we calculated Pearson's r and visually represented correlation using a correlation plot (Figure 2).
"Visualization of Pearson’s correlation between variables with an accompanying
caption and discussion (caption: 1-3 sentences, discussion: 1-3 sentences)

e. Visualization of relationships between variables (not correlation) with an
accompanying caption and discussion (caption: 1-3 sentences, discussion: 1-3
sentences)

f. Creating the null and full model with accompanying discussion (1-3 sentences)
g. Visual and statistical assumption checks for the full model with accompanying
discussion (1-3 sentences)

h. Explanation of any transformation and justification (1-3 sentences)

i. Model construction with visual and statistical assumption checks for three
additional models with justification (i.e. why did you choose those predictor
variables?) (justification: 1-3 sentences per model)

j. Variance inflation factor check for the full model with accompanying discussion
(discussion: 1-3 sentences)

k. Model comparison using the null model, full model, and additional models from part
7g

```{r figure-1}
# visualize missing data 
gg_miss_var(plant) + labs(caption = "Figure 1.")
```

```{r drop-NA}
plant_subset <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_phylls, num_lvs)
```

```{r figure-2}
# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
#returns matrix of numerical values

#create a correlation plot 
corrplot(plant_cor,
         #change shape of cells 
         method = "ellipse",
         addCoef.col = "black")
```

```{r figure-3}
# pair plot can include categorical measurement
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()
```

# Results 
Species:

# Bibliography
