---
title: "Homework 5: how do Sarracenia characteristics predict biomass?"
author: "Katie Miller"
date: "`r Sys.Date()`"
format:
  html_document:
    toc: true
    toc-location: left
    code_folding: true
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

### Introduction Code:
```{r packages}
# loading in necessary packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar)
library(flextable) 
library(car)
library(broom)
library(corrplot)
library(AICcmodavg)
library(GGally)
```

```{r read-data}
# reading in the data, filtering for clarity and relevancy
plant <- read_csv(here("data", "knb-lter-hfr.109.18","hf109-01-sarracenia.csv")) %>% 
  # make column names cleaner
  clean_names() %>% 
  # select columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```


# Methods
Species

a. Sampling methods (these should be summarized from the metadata and cited)

b. Description of data organization and processing (1-3 sentences)

c. We began by creating a visualization of the missing data in our entire dataset (Figure 1). 
Visualization of missing observations with an accompanying caption and discussion
(caption: 1-3 sentences, discussion: 1-3 sentences)

d. In order to examine the relationships between the numerical variables in our dataset, we calculated Pearson's r and visually represented the correlation using a correlation plot (Figure 2). 
"Visualization of Pearson’s correlation between variables with an accompanying
caption and discussion (caption: 1-3 sentences, discussion: 1-3 sentences)

e. (Figure 3) 
Visualization of relationships between variables (not correlation) with an
accompanying caption and discussion (caption: 1-3 sentences, discussion: 1-3
sentences)

f. To determine how species and physiological characters predict biomass, we fit multiple linear models. We started by creating a full model using the entire subset of variables of interest, with total mass ($g$) as the response variable, and the predictor variables being the species of Sarracenia, feedlevel ($g$), specific leaf area ($cm^2/g$), chlorophyll a + b content ($mg/g$), photosynthetic rate ($nmol CO2 * g^-1 * s^-1$), number of phyllodes, and number of pitchers and  phyllodes. The null model does not contain our predictor variables, only the response variable. We tested our full model for normality using Shapiro-Wilk test and for heteroskedasticity using Breusch-Pagan test. Unfortunately, the full model was non-normal & heteroscedastic, therefore did not meet the assumptions for linear regression. To combat this, we performed a mathematical transformation by taking the log of the response variable. 
Creating the null and full model with accompanying discussion (1-3 sentences)

g. Visual and statistical assumption checks for the full model with accompanying
discussion (1-3 sentences)

h. Log transform
Explanation of any transformation and justification (1-3 sentences)

i. Model construction with visual and statistical assumption checks for three
additional models with justification (i.e. why did you choose those predictor
variables?) (justification: 1-3 sentences per model)

j. Variance inflation factor check for the full model with accompanying discussion
(discussion: 1-3 sentences)

k. Model comparison using the null model, full model, and additional models from part
7g


### Methods Code

Figure 1:
```{r figure-1}
# Figure 1
# visualize missing data 
gg_miss_var(plant) + labs(caption = "Figure 1.")


# creating a subset of the data dropping NA values shown in Figure 1
plant_subset <- plant %>%  
  drop_na(sla, chlorophyll, amass, num_phylls, num_lvs)


# Figure 2 
# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
#returns matrix of numerical values


#create a correlation plot 
corrplot(plant_cor,
         #change shape of cells 
         method = "ellipse",
         addCoef.col = "black")


# Figure 3
# pair plot can include categorical measurement
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()

# Full and null models of filtered data

# Assumptions 

# Log transformation

# Assumptions

```


# Results 
species alone can be a good predictor but is not the best because it excludes
(species + sla + feedlevel + chlorophyll + num_lvs)

number of pitchers and phylodes
use MuMIn::dredge()      best: chl, fdl, num_lvs, sla
Species:

# Bibliography
