---
title: "Homework 5"
author: "Katie Miller"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti
execute:
  message: false
  warning: false
editor: visual
---

# Introduction

### Introduction Code:

```{r packages}
# loading in necessary packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar)
library(flextable) 
library(car)
library(broom)
library(corrplot)
library(AICcmodavg)
library(GGally)
```

```{r read-data}
# reading in the data, filtering for clarity and relevancy
plant <- read_csv(here("data", "knb-lter-hfr.109.18","hf109-01-sarracenia.csv")) %>% 
  # make column names cleaner
  clean_names() %>% 
  # select columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```

# Methods


Ellison and Farnsworth (2021) conducted a manipulative feeding experiment with multiple species of Sarracenia to examine the effects of varying nutrient availability on multiple physiological factors. The resulting data was wrangled and uploaded to RStudio. We initially filtered the dataset to extract the variables most relevant to answering our research question: how do Sarracenia characteristics predict biomass? After initial wrangling and filtering, we began our analysis by creating a visualization of the missing data in our entire data subset (Figure 1). This figure brings attention to the missing values for several variables, however, the missingness is relatively low. In order to examine the relationships between the numerical variables in our dataset, we calculated Pearson's r and visually represented the correlation using a correlation plot (Figure 2). This correlation plot allows us to visually represent the possible correlations between our variables. To further understand the relationships between our variables, we created a pair plot of each variable compared against the others (Figure 3). 

To determine how species and physiological characters predict biomass, we fit multiple linear models. We started by creating a full model using the entire subset of variables of interest, with total mass ($g$) as the response variable, and the predictor variables being the species of Sarracenia, feed level ($g$), specific leaf area ($cm^2/g$), chlorophyll a + b content ($mg/g$), photosynthetic rate ($nmol CO2 * g^-1 * s^-1$), number of phyllodes, and number of pitchers and phyllodes. The null model does not contain our predictor variables, only the response variable. We tested our full model for normality using Shapiro-Wilk test and for heteroscedasticity using Breusch-Pagan test. Unfortunately, the full model was non-normal & heteroscedastic, therefore did not meet the assumptions for linear regression (Figure 4). To combat this, we performed a mathematical transformation by taking the log of the response variable. The resulting log transformed model did meet the visual and statistical assumption checks for linear regression by having normally distributed residuals and homoscedastic variance (Figure 5). Using this log model, we set out to determine which predictor variables best explain the response in total mass. 

We found that species alone is a good predictor of total mass. Model 1, representing species as the sole predictor of biomass, conforms visually and statistically to the assumptions of linear regression (Figure 6). We chose this model due to the strong correlation between species and biomass. To create Model 2 we excluded the predictor variable that had the lowest correlation with biomass, which was number of phyllodes. Model 2 also has normally distributed residuals and homoscedastic variance (Figure 7). We hoped that excluding the variable with the weakest correlation would increase the simplicity of our results. Finally, we excluded several poorly correlated variables to create Model 3, including only the more strongly correlated variables of species, specific leaf area, feed level, chlorophyll, and number of pitchers and phyllodes. 
To ensure accuracy of our interpretation, we evaluated multicolinearity by calculating the generalized Variance Inflation Factor and determined that there are no related predictors which inflate the data (Table 1). We then used Akaike Info Criterion (AIC) to compare our constructed models (Table 2).


### Methods Figures & Code


```{r miss-var}
# Figure 1. visualize missing data
gg_miss_var(plant) 

```
**Figure 1. Visualizing variables with missing data in the Sarracinia data subset.** The variables chlorophyll, amass, sla, num_phylls, and num_lvs all have missing data, however, the missingness is relatively low.

```{r subset}
# creating a subset of the data dropping NA values shown in Figure 1
plant_subset <- plant %>%  
  drop_na(sla, chlorophyll, amass, num_phylls, num_lvs)
```

```{r corr-plot}
# correlation plot
# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
#create a correlation plot 
corrplot(plant_cor,
         #change shape of cells 
         method = "ellipse",
         addCoef.col = "black") 
```
**Figure 2. Plot of correlation between variables.** These values were calculated using Pearson's r.

```{r pairs-plot}
# pair plot
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()
```
**Figure 3. Pair plot.** 


```{r full-model}
# Full and null models of filtered data
# null = no predictors
null <- lm(totmass ~ 1, data = plant_subset)
# full = response and all potential predictors 
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)
```

```{r full-vis}
# Full model visual assumptions check
par(mfrow = c(2,2))
plot(full)
```
**Figure 4. Visual check of assumptions for the full model**

```{r full-stats}
# Full model stats assumptions check 
check_normality(full)
check_heteroscedasticity(full)
```

```{r log-transform}
# Log transformation of response variable for null and full models
null_log <- lm(log(totmass) ~ 1, data = plant_subset)
full_log <- lm(log(totmass) ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)

# Assumptions
# Log model visual assumptions check
par(mfrow = c(2,2))
plot(full_log)
```
**Figure 5. Visual check of assumptions for the log model**

```{r log-stats}
# Log model stats assumptions check 
check_normality(full_log)
check_heteroscedasticity(full_log)
```


```{r model1}
# model 1, species predicts biomass?
model1_log <- lm(log(totmass) ~ species, data = plant_subset)

# Assumptions
# model 1 visual assumptions check
par(mfrow = c(2,2))
plot(model1_log)
```
**Figure 6. Visual assumption check for Model 1.**

```{r model1-stat}
# model 1 stats assumptions check 
check_normality(model1_log)
check_heteroscedasticity(model1_log)
```

```{r model2}
# model 2: species + feedlevel + sla + chlorophyll + amass + num_lvs
model2_log <- lm(log(totmass) ~ species + feedlevel + sla + chlorophyll + amass + num_lvs, data = plant_subset)

par(mfrow = c(2,2))
plot(model2_log)
```
**Figure 6. Visual assumption check for Model 2.**

```{r}
# model 2 stats assumptions check 
check_normality(model2_log)
check_heteroscedasticity(model2_log)
```


```{r model3}
# model 3: species + sla + feedlevel + chlorophyll + num_lvs
model3_log <- lm(log(totmass) ~ species + sla + feedlevel + chlorophyll + num_lvs, data = plant_subset)

par(mfrow = c(2,2))
plot(model3_log)
```
**Figure 6. Visual assumption check for Model 3.**

```{r model3-stats}
# model 3 stats assumptions check 
check_normality(model3_log)
check_heteroscedasticity(model3_log)
```


```{r  vif}
# Variance Inflation Factor (VIF) to evaluate multicolinearity
car::vif(full_log)
```
**Table 1. Variance Inflation Factor (VIF).**

```{r aic}
# Akaike Information Criterion (AIC)
MuMIn::model.sel(full_log, null_log, model1_log, model2_log, model3_log)
```
**Table 2. Akaike Information Criterion (AIC).**

# Results

Model 1 showing species alone can be a good predictor but is not the best because it excludes Conversely, Model 2 includes predictor variables with little correlation to the response which negatively effects the accuracy of the overall model. Model 3 is the best model because it is the simplest as determined by AIC Akaike Info Criterion (AIC) from Table x values show that Model 3 is the simplest model that explains the most variance. (species + sla + feedlevel + chlorophyll + num_lvs)

number of pitchers and phylodes use MuMIn::dredge() best: chl, fdl, num_lvs, sla Species:

We found the best predictor variable of total mass of a Sarracina plant to be the total number of pitchers and phyllodes it possesses. Biologically, this makes sense because the more pitchers and phyllodes would mean a larger plant, and therefore larger total biomass.

### Results Code

```{r}
summary(model3_log)
```

```{r}
model_pred <- ggpredict(model3_log, terms = "num_lvs", back.transform = TRUE)

model_pred
```

```{r}
plot(ggpredict(model3_log, terms = "num_lvs", back.transform = TRUE), add.data = TRUE)
```

# Bibliography
