---
title: "Homework 5"
author: "Katie Miller"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti
execute:
  message: false
  warning: false
editor: visual
---

# Introduction

### Introduction Code:
```{r packages}
# loading in necessary packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar)
library(flextable) 
library(car)
library(broom)
library(corrplot)
library(AICcmodavg)
library(GGally)
```

```{r read-data}
# reading in the data, filtering for clarity and relevancy
plant <- read_csv(here("data", "knb-lter-hfr.109.18","hf109-01-sarracenia.csv")) %>% 
  # make column names cleaner
  clean_names() %>% 
  # select columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```


# Methods
Species

a. Sampling methods (these should be summarized from the metadata and cited)

b. Description of data organization and processing (1-3 sentences)

c. We began by creating a visualization of the missing data in our entire dataset (Figure 1). 
Visualization of missing observations with an accompanying caption and discussion
(caption: 1-3 sentences, discussion: 1-3 sentences)

d. In order to examine the relationships between the numerical variables in our dataset, we calculated Pearson's r and visually represented the correlation using a correlation plot (Figure 2). 
"Visualization of Pearson’s correlation between variables with an accompanying
caption and discussion (caption: 1-3 sentences, discussion: 1-3 sentences)

e. (Figure 3) 
Visualization of relationships between variables (not correlation) with an
accompanying caption and discussion (caption: 1-3 sentences, discussion: 1-3
sentences)

f. To determine how species and physiological characters predict biomass, we fit multiple linear models. We started by creating a full model using the entire subset of variables of interest, with total mass ($g$) as the response variable, and the predictor variables being the species of Sarracenia, feed level ($g$), specific leaf area ($cm^2/g$), chlorophyll a + b content ($mg/g$), photosynthetic rate ($nmol CO2 * g^-1 * s^-1$), number of phyllodes, and number of pitchers and  phyllodes. The null model does not contain our predictor variables, only the response variable. We tested our full model for normality using Shapiro-Wilk test and for heteroskedasticity using Breusch-Pagan test. Unfortunately, the full model was non-normal & heteroscedastic, therefore did not meet the assumptions for linear regression. To combat this, we performed a mathematical transformation by taking the log of the response variable. 
Creating the null and full model with accompanying discussion (1-3 sentences)

g. Visual and statistical assumption checks for the full model with accompanying
discussion (1-3 sentences)

h. Log transform
Explanation of any transformation and justification (1-3 sentences)

i. Using this log model, we set out to determine which set of predictor variables best explains the response in total mass. We found that species alone is a good predictor of total mass. 
Figure 4: model 1, species (not enough)


Assumption

Figure 5: model 2, 

To create the second model, we removed the predictor variable with the low correlation with biomass, which was the  returns a viable model. Looking at a model with relatively many predictor variables 
Assumptions

Figure 6: model 3, goldilocks

Model construction with visual and statistical assumption checks for three
additional models with justification (i.e. why did you choose those predictor
variables?) (justification: 1-3 sentences per model)

j. We evaluated multicolinearity by calculating generalized variance inflation factor and determined that there are no related predictors which inflate the data.
Variance inflation factor check for the full model with accompanying discussion
(discussion: 1-3 sentences)

k. Model comparison using the null model, full model, and additional models from part
7g


### Methods Code

Figure 1:
```{r methods-code}
# Figure 1: visualize missing data
gg_miss_var(plant) + labs(caption = "Figure 1.")


# creating a subset of the data dropping NA values shown in Figure 1
plant_subset <- plant %>%  
  drop_na(sla, chlorophyll, amass, num_phylls, num_lvs)


# Figure 2: correlation plot
# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
#create a correlation plot 
corrplot(plant_cor,
         #change shape of cells 
         method = "ellipse",
         addCoef.col = "black",
         title = "Figure 2.") 


# Figure 3: pair plot
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()

# Full and null models of filtered data
# null = no predictors
null <- lm(totmass ~ 1, data = plant_subset)
# full = response and all potential predictors 
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)

# Full model visual assumptions check
par(mfrow = c(2,2))
plot(full, caption = "Table x. Visual check of assumptions for the full model")

# Full model stats assumptions check 
check_normality(full)
check_heteroscedasticity(full)

# Log transformation of response variable for null and full models
null_log <- lm(log(totmass) ~ 1, data = plant_subset)
full_log <- lm(log(totmass) ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)

# Assumptions
# Log model visual assumptions check
par(mfrow = c(2,2))
plot(full_log, caption = "Table x. Visual check of assumptions for the log model")

# Log model stats assumptions check 
check_normality(full_log)
check_heteroscedasticity(full_log)

# Figure 4: model 1 
# species predicts biomass?
model1_log <- lm(log(totmass) ~ species, data = plant_subset)

# Assumptions
# model 1 visual assumptions check
par(mfrow = c(2,2))
plot(model1_log, caption = "Table x. Visual check of assumptions for the log model")

# model stats assumptions check 
check_normality(model1_log)
check_heteroscedasticity(model1_log)

# Figure 5: model 2

# Assumptions

# Figure 6: model 3
# species predicts biomass?
model3_log <- lm(log(totmass) ~ species + sla + feedlevel + chlorophyll + num_lvs, data = plant_subset)

plot(model3_log)

# model stats assumptions check 
check_normality(model3_log)
check_heteroscedasticity(model3_log)


# Variance Inflation Factor (VIF) to evaluate multicolinearity
car::vif(full_log, caption = "Table x.")

# Akaike Info Criterion (AIC)
# MuMIn::model.sel(full_log, null_log, model1_log, model2_log, model3_log)
```


# Results 
species alone can be a good predictor but is not the best because it excludes
(species + sla + feedlevel + chlorophyll + num_lvs)

number of pitchers and phylodes
use MuMIn::dredge()      best: chl, fdl, num_lvs, sla
Species:

# Bibliography
