---
title: "HW5"
author: "Katie"
date: "2023"
output:
  html_document:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti
execute:
  message: false
  warning: false
---

```{r}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```

read in data
```{r read-data}
plant <- read_csv(here("data", "knb-lter-hfr.109.18","hf109-01-sarracenia.csv")) %>% 
  # make column names cleaner
  clean_names() %>% 
  # select columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)

```

```{r miss-vis}
gg_miss_var(plant)
```

subset dropping NA
```{r}
plant_subset <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_phylls, num_lvs)
```

create a correlation plot

"to determine the relationships between numerical variables in our dataset, we calculated pearsons r and visually represented correlation using a correlation plot"
```{r correlation-plot}
# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
#returns matrix of numerical values

#create a correlation plot 
corrplot(plant_cor,
         #change shape of cells 
         method = "ellipse",
         addCoef.col = "black")

#don't include two predictors (eg mass) because can conflate r^2



```

create a plot of each variable compared against others
```{r}
#pair plot can include categorical measurement
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()
```

starting regression here:
"to determine how species and physiological characters predict bimass, we fit multiple linear models"
```{r}
# null = no predictors
null <- lm(totmass ~ 1, data = plant_subset)
# full = response and all potential predictors 
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)
```

we visually asses normality and homoskedasticity of residuals using diagnostic plots for full model 
```{r}
par(mfrow = c(2,2))
plot(full)
```

We also tested for normality using Shapiro-Wilk test (null: variable of interest (ie residuals) are normally distributed). 
We tested for heteroskedasticity using Breusch-Pagan test (null: variable of interest has constant variance)
```{r}
check_normality(full)
check_heteroscedasticity(full)
```
Non-normality & Heteroscedasticity;
Assumptions of linear regression are not met!

mathematical transformation can make data betta
(only transform response, predictors can stay on the same scale)
```{r}
# log transform response
null_log <- lm(log(totmass) ~ 1, data = plant_subset)
full_log <- lm(log(totmass) ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)

plot(full_log)
```


check normality of full model with log transform
```{r}
# do any combo of predictors inflate the data (because they're related)
check_normality(full_log)
check_heteroscedasticity(full_log)
```


they're good now! meeting assumption on linear regression
all interpretation will be based on a transformed response 


Evaluate multicolinearity
```{r}
car::vif(full_log)
```
We evaluated multicolinearity by calculating generalized variance inflation factor and determined that there are no related predictors which inflate the data.

try some models 
addressing question: what set of predictor variables best explains the response

```{r}
# species predicts biomass?

model2_log <- lm(log(totmass) ~ species, data = plant_subset)
```

```{r}
# check assumptions for model 

plot(model2_log)
# clumps = predictor categorical, all good

check_normality(model2_log)
check_heteroscedasticity(model2_log)
# conforms to assumptions of linear regression!
```

compare models using Akaike Info Criterion (AIC) values:
(looks for the simplest model that explains the most variance)
```{r}
AICc(full_log)
AICc(model2_log)    
AICc(null_log)
AICc(model3_log)
# least complex model that best predicts the response is full model (lowest value)
```

```{r}
MuMIn::AICc(full_log, model2_log, null_log)
# if you want to do them all at once 

MuMIn::model.sel(full_log, model2_log, null_log)
# more info, yay!
```

we compared models using AIC and chose the model with the lowest value which was..


# Results 

we found that the __ model including __ __ __ predictors best predicted __ (model summary). 

```{r}
summary(full_log)

```

```{r}
table <- tidy(full_log, conf.int = TRUE) %>% 
  #change the p-value numbers if they tiny (>0.001)
  # change est, SE, and t-stat to round to 
  #using mutate
  #make into flex table 
  flextable() %>% 
  #fit it to viewer
  autofit()
  

table
```

back transform estimates by exponentiating (because we took log)
```{r}
model_pred <- ggpredict(full_log, terms = "species", back.transform = TRUE)

# reporting results in original variable is fine if you tell ppl that 

# "log transformed to meet assumptions, back transformed for ease of communication of results"
```

```{r}
model_pred
```

all else held constant, we would expect total mass to be around predicted +/- CI. 

```{r}
# plot to visualize relationships in data (can choose significant from summary(full_log))
plot(ggpredict(full_log, terms = "species", back.transform = TRUE), add.data = TRUE)

plot(ggpredict(full_log, terms = "chlorophyll", back.transform = TRUE), add.data = TRUE)

# choose which figure best represents the main take away of your analysis
```


# different types of ANOVA 

type 1: order matter; species > feed level > etc
type 2: order doesnt mateer but there may be interaction
type 3: order and interaction matters

```{r}

```



# for my HW
possible predictors of biomass are
species + feedlevel + sla + çhlorophyll + amass + num_lvs + num_phylls

```{r}
plot(ggpredict(full_log, terms = "amass", back.transform = TRUE), add.data = TRUE)
```

```{r}
model3_log <- lm(log(totmass) ~ chlorophyll, data = plant_subset)

plot(model3_log)
# clumps = predictor categorical, all good

```

```{r}
check_normality(model3_log)
check_heteroscedasticity(model3_log)
```

```{r}
MuMIn::AICc(full_log, model2_log, null_log, model3_log)
```

* sla, 275
** amass, 295
*** species,158
